---
title: "Analysis of descriptive statistics supporting impact evaluation of the Frailty Improvement Collaborative"
format: 
  html: 
    toc: true
    self-contained: true
    html-math-method: katex

editor: visual
---

```{r, include = FALSE}

rm(list = ls())
library(tidyverse)
library(gt)
library(zoo)

sicb_lookup <- read_csv(
  file = "../2b_Wrangled_Data/sicb_lookup.csv") |> 
  subset(frail_inds_across_all_gp > 10000)

sicb_yearmon_outcomes <- read_csv(
  file = "../2b_Wrangled_Data/sicb_yearmon_outcomes.csv") |> 
  subset(sicb_code %in% sicb_lookup$sicb_code) |> 
  mutate(YearMon = zoo::as.yearmon(YearMon)) |> 
  subset(YearMon >= "Apr 2019")

sicb_lookup <- left_join(
  x = sicb_lookup, 
  y = unique(select(sicb_yearmon_outcomes, sicb_code, place_name)))

# set max date as four months before the latest month included in the dataset
max_date = max(sicb_yearmon_outcomes$YearMon) - (4/12)
```

## Overview

### Introduction

This document provides a descriptive overview of the key metrics related to the impact evaluation of the Frailty Improvement Collaborative.

### Identifying places

All outcomes were constructed by linking the outcomes of individual patients back to their GP practices, with GP practices then allocated to geographical places, in order to construct outcomes for each place.

Interventions happened at the level of 'place'. This aligned with sub-ICB locations as used internally by ICBs. However, the available national datasets use a definition and boundary of sub-ICB locations which do not necessarily match the internal definition used by that ICB. For example, in the available national datasets the Black Country ICB has a single sub-ICB location which covers the whole ICB - and is not broken down any further. For this reason it was not possible to directly use the most recent sub-ICB locations as defined in national datasets to identify the discovery places (e.g. Wolverhampton in the case of the Black Country).

The CCG boundaries as defined in the financial year 2017/18 were instead used to identify places. In many cases the 2017/18 CCG areas map directly on to current sub-ICB locations. However in some cases, they provided breakdowns better aligned with the definition of places as used by this programme. For example, the 2017/18 CCG boundaries break down the Black Country ICB into multiple sub-locations, one of which was Wolverhampton (which was one of the places which received the intervention).

In the case of Gloucestershire, the defined 'place' was a PCN (North and South Gloucester). For the purpose of this analysis, in order to make results comparable to other areas, the affected area was taken to be the Integrated Locality Partnership (ILP) that the PCN sat within (Gloucester). Even though it includes a slightly larger area (taking in other parts of Gloucester) it remains one of the smaller places.

Using the 2017/18 CCG boundaries, there would be 207 places, although due to Gloucestershire being broken down into six ILPs (rather than the one sub-ICB in Gloucestershire ICB), the country is instead broken down into 212 places. Seven ICBs introduced the programme, although in some ICBs, two places were listed as taking part in the programme, resulting in a total of ten places which took part in the programme.

A small number of GPs were excluded due to the process of allocating GPs to places - for example, they were created since 2018 and were allocated to sub-ICB codes which do not have a direct counterpart in the 2017/18 CCG boundaries. Some GPs were also excluded due to not aligning directly with a geographic region (i.e. the GP code is exclusively used for military service personnel or for homeless individuals across an ICB).

### Definition of frailty

For the purposes of this analysis, patients were defined as frail if they had an intermediate risk or high risk based on their [Hospital Frailty Score](https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(18)30668-8/fulltext) (HFRS), where the HFRS was constructed for each patient at any given point in time constructed using their diagnoses within the preceding two years.

### Variables used

The available data extended to April 2018, although outcomes for people living with frailty were misleadingly low in the early months because there was not enough preceding data to accurately identify people living with frailty on former diagnoses. For this reason, all outcomes are shown from April 2019, meaning there is at least a full year's worth of data with which to identify which patients are frail.

The following descriptive variables are included here:

-   Frail population size,

-   Frailty-weighted deprivation score,

-   Frailty-weighted rural index score.

The following outcomes are included here as potential outcomes for an impact evaluation:

-   Reduction in hospital admissions for people living with frailty,

-   Reduced average length of stay for people living with frailty,

-   Fewer total bed days used by people living with frailty,

-   Reduction in delays to discharge and faster transitions to intermediate care,

-   Reduced attendances to Emergency Departments for people living with frailty,

-   Fewer hospital re-admissions within 28 days of discharge,

-   Reduction in hospital conveyances for people living with frailty,

-   Proportionally more discharges to usual place of residence.

## Descriptive variables

### Frail population

The frail population was determined by counting all of the individuals within a given place who had a HFRS of five or higher at any point during the period covered.

```{r, include = FALSE}

frail_inds <- signif(sum(7724812),3)
frail_inds2 <- signif(sum(sicb_lookup$frail_inds_across_all_gp), 3)

mean_frails <- signif(mean(sicb_lookup$frail_inds_across_all_gp), 3)
max_frails <- signif(max(sicb_lookup$frail_inds_across_all_gp), 3)
min_frails <- signif(min(sicb_lookup$frail_inds_across_all_gp), 3)

```

Across all included places, across the entire period covered (from April 2019 to October 2025), there were `r format(frail_inds, scientific = FALSE, big.mark = ",")` individuals who were identified as being frail at any point. Note that within the dataset used for analysis here, it appears to be `r format(frail_inds2, scientific = FALSE, big.mark = ",")` individuals, due to double-counting where people living with frailty have changed GP. This issue should not cause any issues in later outcomes or the eventual analysis, as all outcomes focus on the number of attendances by people living with frailty, rather than a count of the people living with frailty themselves, and so should not be affected by individuals moving GP.

The average place has a mean of `r format(mean_frails, scientific = FALSE, big.mark = ",")` people living with frailty across the whole period, although this varies from a maximum of `r format(max_frails, scientific = FALSE, big.mark = ",")` to a minimum of `r format(min_frails, scientific = FALSE, big.mark = ",")` .

The magnitude of the frail population for each of the intervention places is shown in the table below.

```{r, include = FALSE}

frail_pop_tbl <- sicb_lookup |> 
  subset(is.na(place_name) == FALSE) |> 
  transmute('Place' = place_name, 
            'Frail population' = frail_inds_across_all_gp) |> 
  mutate('Frail population relative to average (%)' = round(100*(`Frail population`/mean_frails), 1))

```

```{r, echo = FALSE}

frail_pop_tbl |> 
  gt()
```

### Frailty-weighted deprivation score

```{r, include = FALSE}

deprivation <- round(sum(sicb_lookup$depr_frail_inds)/sum(sicb_lookup$frail_inds_across_all_gp), 3)

sicb_lookup <- sicb_lookup |> 
  mutate(depr_prop = depr_frail_inds/frail_inds_across_all_gp)

min_depr <- round(min(sicb_lookup$depr_prop), 3)
max_depr <- round(max(sicb_lookup$depr_prop), 3)
```

Places were given a frailty-weighted deprivation score by taking the proportion of people living with frailty within a place who live in the 20% most deprived Lower Super Output Areas (LSOAs) in the country (using the 2025 Index of Multiple Deprivation).

Across all places, `r format(deprivation*100)`% of people living with frailty lived in deprived LSOAs, reflecting the fact that people living with frailty are slightly more likely to be deprived than the national average (20%). However this varies substantially by place, from `r format(min_depr*100)`% at the lowest to `r format(max_depr*100)`% at the highest.

The frailty-weighted deprivation score of each of the intervention places is shown below.

```{r, include = FALSE}

depr_tbl <- sicb_lookup |> 
  subset(is.na(place_name) == FALSE) |> 
  transmute('Place' = place_name, 
            'Frailty-weighted deprivation score (%)' = round(100 * depr_prop, 1))

```

```{r, echo = FALSE}

depr_tbl |> 
  gt()
```

### Frailty-weighted rural index score

```{r, include = FALSE}

rurality <- round(sum(sicb_lookup$rural_frail_inds)/sum(sicb_lookup$frail_inds_across_all_gp), 3)

sicb_lookup <- sicb_lookup |> 
  mutate(rural_prop = rural_frail_inds/frail_inds_across_all_gp)

min_rural <- round(min(sicb_lookup$rural_prop), 3)
max_rural <- round(max(sicb_lookup$rural_prop), 3)

```

Places were given a frailty-weighted rural index score by taking the proportion of people living with frailty within a place who live in a rural Lower Super Output Areas (LSOAs).

Across all places, `r format(rurality*100)`% of people living with frailty lived in rural LSOAs, which is approximately in line with the national average (17%). However this varies substantially by place, from `r format(min_rural*100)`% at the lowest to `r format(max_rural*100)`% at the highest.

The frailty-weighted rural index score of each of the intervention places is shown below.

```{r, include = FALSE}

rural_tbl <- sicb_lookup |> 
  subset(is.na(place_name) == FALSE) |> 
  transmute('Place' = place_name, 
            'Frailty-weighted rural index score (%)' = round(100 * rural_prop, 1))
```

```{r, echo = FALSE}

rural_tbl |> 
   gt()
```

## Potential impact evaluation outcomes

### Hospital admissions for people living with frailty

Hospital admissions by people living with frailty is defined as the number of unplanned admissions to hospital by patients with a HFRS of five or over at the time of admission.

```{r, include = FALSE}

sum_apc_admiss <- signif(sum(subset(sicb_yearmon_outcomes, YearMon <= max_date)$apc_admissions),3)
mean_apc_admiss <- signif(mean(subset(sicb_yearmon_outcomes, YearMon <= max_date)$apc_admissions),3)

place_apc_admissions <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name) |> 
  summarise(mean_apc_admissions = round(mean(apc_admissions), 1)) |> 
  ungroup() |> 
  transmute(Place = place_name, 
            'Average hospital admissions' = mean_apc_admissions)

apc_admissions_plot <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name, YearMon) |> 
  summarise(apc_admissions = mean(apc_admissions)) |> 
  ungroup() |> 
  mutate(group = ifelse(place_name == "Average of control places", 
                        yes = "Control", 
                        no = "Intervention")) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = apc_admissions, 
                       col = place_name)) + 
  geom_line() + 
  scale_y_continuous(name = "Frail hospital admissions", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  scale_colour_discrete(name = "Place")  + 
  facet_wrap(group~., 
             nrow = 2, 
             ncol = 1) +
  theme_minimal() + 
  theme(legend.position = "bottom")

  
```

Across the whole period, there were `r format(sum_apc_admiss, scientific = FALSE, big.mark = ",")` hospital admissions by people living with frailty across all places, which equated to an average of `r format(mean_apc_admiss, scientific = FALSE, big.mark = ",")` per place per month.

The average number of hospital admissions by people living with frailty per month, broken down by place, is shown in the table below.

```{r, echo = FALSE}

place_apc_admissions |> 
  gt()
```

The national trend for hospital admissions by people living with frailty per month is shown on the chart below.

```{r, echo = FALSE}

sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  group_by(YearMon) |> 
  summarise(apc_admissions = sum(apc_admissions)) |> 
  ungroup() |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = apc_admissions)) + 
  geom_line() + 
  theme_minimal() + 
  scale_y_continuous(name = "Frail hospital admissions", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date")

```

The place level trend for hospital admissions by people living with frailty per month is shown on the chart below (relative to the average for all the control places).

```{r, echo = FALSE}

plotly::ggplotly(apc_admissions_plot)

```

### Hospital re-admissions for people living with frailty within 28 days of discharge

Hospital readmissions for people living with frailty within 28 days of discharge is a count of the total number of unplanned admissions to hospital by patients with a HFRS of five or over at the time of admissions who also had a discharge within the preceding 28 days. It is worth noting that this definition also excludes patients with various conditions where treatment can involve repeated admission and discharge.

```{r, include = FALSE}

sum_apc_readmiss <- signif(sum(subset(sicb_yearmon_outcomes, YearMon <= max_date)$apc_readmissions),3)
mean_apc_readmiss <- signif(mean(subset(sicb_yearmon_outcomes, YearMon <= max_date)$apc_readmissions),3)

place_apc_readmissions <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name) |> 
  summarise(mean_apc_readmissions = round(mean(apc_readmissions), 1)) |> 
  ungroup() |> 
  transmute(Place = place_name, 
            'Average hospital readmissions' = mean_apc_readmissions)

apc_readmissions_plot <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name, YearMon) |> 
  summarise(apc_readmissions = mean(apc_readmissions)) |> 
  ungroup() |> 
  mutate(group = ifelse(place_name == "Average of control places", 
                        yes = "Control", 
                        no = "Intervention")) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = apc_readmissions, 
                       col = place_name)) + 
  geom_line() + 
  scale_y_continuous(name = "Frail hospital readmissions", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  scale_colour_discrete(name = "Place") + 
    facet_wrap(group~.,               
               nrow = 2,               
               ncol = 1) +
  theme_minimal()  + 
  theme(legend.position = "bottom")
```

Across the whole period, there were `r format(sum_apc_readmiss, scientific = FALSE, big.mark = ",")` hospital readmissions by people living with frailty across all places, which equated to an average of `r format(mean_apc_readmiss, scientific = FALSE, big.mark = ",")` per place per month.

The average number of hospital readmissions by people living with frailty per month, broken down by place, is shown in the table below.

```{r, echo = FALSE}

place_apc_readmissions |> 
  gt()

```

The national trend for hospital readmissions by people living with frailty is shown on the chart below.

```{r, echo = FALSE}

sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  group_by(YearMon) |> 
  summarise(apc_readmissions = sum(apc_readmissions)) |> 
  ungroup() |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = apc_readmissions)) + 
  geom_line() + 
  theme_minimal() + 
  scale_y_continuous(name = "Frail hospital readmissions", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date")
```

The place level trend for hospital readmissions by people living with frailty per month is shown on the chart below (relative to the average for all the control places).

```{r, echo = FALSE}

plotly::ggplotly(apc_readmissions_plot)
```

### Attendances to Emergency Departments for people living with frailty

Attendances to Emergency Departments (ED) for people living with frailty is defined as the number of unplanned attendances to ED by patients with a HFRS of five or over at the time of attendance.

```{r, include = FALSE}

sum_ed_admiss <- signif(sum(subset(sicb_yearmon_outcomes, YearMon <= max_date & YearMon >= "Apr 2019")$ed_admissions),3)
mean_ed_admiss <- signif(mean(subset(sicb_yearmon_outcomes, YearMon <= max_date & YearMon >= "Apr 2019")$ed_admissions),3)

place_ed_admissions <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date & YearMon >= "Apr 2019") |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name) |> 
  summarise(mean_ed_admissions = round(mean(ed_admissions), 1)) |> 
  ungroup() |> 
  transmute(Place = place_name, 
            'Average ED admissions' = mean_ed_admissions)

ed_admissions_plot <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date & YearMon >= "Apr 2019") |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name, YearMon) |> 
  summarise(ed_admissions = mean(ed_admissions)) |> 
  ungroup() |> 
  mutate(group = ifelse(place_name == "Average of control places", 
                        yes = "Control", 
                        no = "Intervention")) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = ed_admissions, 
                       col = place_name)) + 
  geom_line() + 
  scale_y_continuous(name = "Frail ED admissions", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  scale_colour_discrete(name = "Place") + 
    facet_wrap(group~.,               
               nrow = 2,               
               ncol = 1) +
  theme_minimal()  + 
  theme(legend.position = "bottom")
```

Across the whole period, there were `r format(sum_ed_admiss, scientific = FALSE, big.mark = ",")` ED admissions by people living with frailty across all places, which equated to an average of `r format(mean_ed_admiss, scientific = FALSE, big.mark = ",")` per place per month.

The average ED admissions by people living with frailty per month for each place is shown in the table below.

```{r, echo = FALSE}

place_ed_admissions |>    
  gt()
```

The national trend for ED admissions by people living with frailty per month is shown on the chart below.

```{r, echo = FALSE}

sicb_yearmon_outcomes |>    
  subset(YearMon <= max_date & YearMon >= "Apr 2019") |>    
  group_by(YearMon) |>    
  summarise(ed_admissions = sum(ed_admissions)) |>    
  ungroup() |>    
  ggplot(mapping = aes(x = YearMon,                         
                       y = ed_admissions)) +    
  geom_line() +    
  theme_minimal() +    
  scale_y_continuous(
    name = "Frail  ED admissions",                       
    labels = scales::label_comma()) +    
  scale_x_yearmon(name = "Date") 
```

The place level trend for ED admissions by people living with frailty per month is shown on the chart below (relative to the average for all the control places)

```{r, echo = FALSE}

plotly::ggplotly(ed_admissions_plot)
```

### Total bed days used by people living with frailty

Total bed days used by people living with frailty is the sum of completed days in hospital by patients with a HFRS of five or higher at the time of admissions.

```{r, include = FALSE}


sum_apc_hospital_days <- signif(sum(subset(sicb_yearmon_outcomes, YearMon <= max_date)$apc_hospital_days),3)
mean_apc_hospital_days <- signif(mean(subset(sicb_yearmon_outcomes, YearMon <= max_date)$ed_admissions),3)

place_apc_hospital_days <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name) |> 
  summarise(mean_apc_hospital_days = round(mean(apc_hospital_days), 1)) |> 
  ungroup() |> 
  transmute(Place = place_name, 
            'Monthly average of total hospital days' = mean_apc_hospital_days)

apc_hospital_days_plot <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name, YearMon) |> 
  summarise(apc_hospital_days = mean(apc_hospital_days)) |> 
  ungroup() |> 
  mutate(group = ifelse(place_name == "Average of control places", 
                        yes = "Control", 
                        no = "Intervention")) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = apc_hospital_days, 
                       col = place_name)) + 
  geom_line() + 
  scale_y_continuous(name = "Frail hospital days", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  scale_colour_discrete(name = "Place") + 
    facet_wrap(group~.,               
               nrow = 2,               
               ncol = 1) +
  theme_minimal()  + 
  theme(legend.position = "bottom")

```

Across the whole period, there were `r format(sum_apc_hospital_days, scientific = FALSE, big.mark = ",")` hospital days by people living with frailty across all places, which equated to an average of `r format(mean_apc_hospital_days, scientific = FALSE, big.mark = ",")` per place per month.

The average total number of hospital days by people living with frailty per month for each place is shown in the table below.

```{r, echo = FALSE}

place_apc_hospital_days |>
  gt()
```

The national trend for total hospital days by people living with frailty per month is shown on the chart below.

```{r, echo = FALSE}

sicb_yearmon_outcomes |>    
  subset(YearMon <= max_date) |>    
  group_by(YearMon) |>    
  summarise(apc_hospital_days = sum(apc_hospital_days)) |>    
  ungroup() |>    
  ggplot(mapping = aes(x = YearMon,                         
                       y = apc_hospital_days)) +    
  geom_line() +    
  theme_minimal() +    
  scale_y_continuous(
    name = "Frail hospital days",                       
    labels = scales::label_comma()) +    
  scale_x_yearmon(name = "Date") 
```

The place level trend for total hospital days by people living with frailty per month is shown on the chart below (relative to the average for all the control places).

```{r, echo = FALSE}

plotly::ggplotly(apc_hospital_days_plot)

```

### Average length of stay for people living with frailty

The average length of stay for people living with frailty is the total number of completed days in hospital by patients with a HFRS of five or higher on admission, divided by the number of admissions to hospital by people with a HFRS of five or higher on admission.

```{r, include = FALSE}


mean_apc_los <- sum_apc_hospital_days/sum_apc_admiss

place_apc_los <- sicb_yearmon_outcomes |>    
  subset(YearMon <= max_date) |>    
  mutate(place_name = ifelse(test = is.na(place_name),                              
                             yes = "Average of control places",             
                             no = place_name)) |>    
  group_by(place_name) |>    
  summarise(sum_apc_hospital_days = sum(apc_hospital_days), 
            sum_apc_admiss= sum(apc_admissions)) |>    
  ungroup() |>    
  transmute(Place = place_name,              
            'Average hospital length of stay'= round(
              sum_apc_hospital_days/sum_apc_admiss, 2)) 

apc_los_plot <- sicb_yearmon_outcomes |>    
  subset(YearMon <= max_date) |>   
  mutate(place_name = ifelse(test = is.na(place_name),          
                             yes = "Average of control places",      
                             no = place_name)) |>    
  group_by(place_name, YearMon) |>    
  summarise(sum_apc_hospital_days = sum(apc_hospital_days), 
            sum_apc_admiss= sum(apc_admissions)) |>    
  ungroup() |>    
  mutate(group = ifelse(place_name == "Average of control places",              
                        yes = "Control",                          
                        no = "Intervention")) |>    
  ggplot(mapping = aes(x = YearMon,                         
                       y = sum_apc_hospital_days/sum_apc_admiss,                    
                       col = place_name)) +  
  geom_line() +    
  scale_y_continuous(name = "Frail hospital length of stay",    
                     labels = scales::label_comma()) +    
  scale_x_yearmon(name = "Date") +    
  scale_colour_discrete(name = "Place") +    
    facet_wrap(group~.,               
               nrow = 2,              
               ncol = 1) +   
  theme_minimal()  + 
  theme(legend.position = "bottom") 
```

Across the whole period, across all places, the average hospital length of stay for people living with frailty was `r format(mean_apc_los, scientific = FALSE, big.mark = ",")`.

The average hospital length of stay by people living with frailty for each place is shown in the table below.

```{r, echo = FALSE}

place_apc_los |>   
  gt()
```

The national trend for average hospital length of stay by people living with frailty per month is shown on the chart below.

```{r, echo=FALSE}

sicb_yearmon_outcomes |>    
  subset(YearMon <= max_date) |>    
  mutate(place_name = ifelse(test = is.na(place_name),                              
                             yes = "Average of control places",             
                             no = place_name)) |>    
  group_by(YearMon) |>    
  summarise(sum_apc_hospital_days = sum(apc_hospital_days), 
            sum_apc_admiss= sum(apc_admissions)) |>    
  ungroup() |>    
  transmute(YearMon = YearMon, 
            apc_los = round(
              sum_apc_hospital_days/sum_apc_admiss, 10))  |>       
  ggplot(mapping = aes(x = YearMon,                
                       y = apc_los)) +       
  geom_line() +       
  theme_minimal() +       
  scale_y_continuous(
    name = "Frail hospital length of stay",                            
    labels = scales::label_comma()) +       
  scale_x_yearmon(name = "Date") 
```

The place level trend for average hospital length of stay by people living with frailty per month is shown on the chart below (relative to the average for all the control places).

```{r,echo=FALSE}

plotly::ggplotly(apc_los_plot) 
```

### Hospital conveyances for people living with frailty

Hospital conveyances for people living with frailty was defined as the number of ambulance conveyances to ED for patients with a HFRS of five or higher at attendance, as a proportion of the total number of attendances to ED by patients with a HFRS or five or higher at attendance.

```{r, include=FALSE}

sum_ed_amb_admiss <- signif(sum(subset(sicb_yearmon_outcomes, YearMon <= max_date & YearMon >= "Apr 2019")$ed_admissions_by_ambulance),3)

mean_ed_admiss <- signif(mean(subset(sicb_yearmon_outcomes, YearMon <= max_date & YearMon >= "Apr 2019")$ed_admissions),3)

place_ed_amb_admissions <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date & YearMon >= "Apr 2019") |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name) |> 
  summarise(ed_admissions = mean(ed_admissions), 
            ed_amb_admissions = mean(ed_admissions_by_ambulance)) |> 
  ungroup() |> 
  transmute(Place = place_name, 
            'Average ED attendances by ambulance' = signif(
              (ed_amb_admissions), 3),
            'Average ED attendances' = signif(
              (ed_admissions), 3),
            'ED attendances by ambulance (%)' = round(
              100*(ed_amb_admissions/ed_admissions), 1))

ed_amb_admissions_plot <- sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date & YearMon >= "Apr 2019") |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name, YearMon) |> 
  summarise(ed_admissions = sum(ed_admissions), 
            ed_amb_admissions = sum(ed_admissions_by_ambulance)) |> 
  ungroup() |> 
  mutate(group = ifelse(place_name == "Average of control places", 
                        yes = "Control", 
                        no = "Intervention")) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = 100*(ed_amb_admissions/ed_admissions), 
                       col = place_name)) + 
  geom_line() + 
  scale_y_continuous(name = "Frail ED ambulance attendances (%)", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  scale_colour_discrete(name = "Place") + 
    facet_wrap(group~.,               
               nrow = 2,               
               ncol = 1) +
  theme_minimal()  + 
  theme(legend.position = "bottom")

```

Across the whole period, there were `r format(sum_ed_admiss, scientific = FALSE, big.mark = ",")` ED attendances by people living with frailty across all places, of which `r format(sum_ed_amb_admiss, scientific = FALSE, big.mark = ",")` were by ambulance (or `r format(100*(sum_ed_amb_admiss/sum_ed_admiss), scientific = FALSE, big.mark = ",")` %)

The average frail ED ambulance admissions over the whole period for each intervention place is shown in the table below.

```{r, echo=FALSE}

place_ed_amb_admissions |> 
  gt()

```

The national trend for frail ED ambulance admissions as a percentage per month is shown on the chart below.

```{r, echo=FALSE}

sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date & YearMon >= "Apr 2019") |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(YearMon) |> 
  summarise(ed_admissions = sum(ed_admissions), 
            ed_amb_admissions = sum(ed_admissions_by_ambulance)) |> 
  ungroup() |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = 100*(ed_amb_admissions/ed_admissions))) + 
  geom_line() + 
  scale_y_continuous(name = "Frail ED ambulance attendances (%)", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  theme_minimal() 

```

The place level trend for frail ED attendances by ambulance (as a percentage) per month is shown on the chart below (relative to the average for all the control places).

```{r, echo=FALSE}

plotly::ggplotly(ed_amb_admissions_plot)
```

### Delays to discharge and faster transitions to intermediate care for people living with frailty

Delays in discharge for people living with frailty was defined as the number of patients with a HFRS of five or higher on admission, who were discharged after their Discharge Ready Date (DRD), as a proportion of all the discharged patients with a HFRS of five or higher on admissions.

Hospital discharges are grouped into five categories, in relation to the patient's discharge ready date (DRD):

-   discharged on the same day as admission

-   discharged after the DRD

-   discharged on the DRD

-   discharged before the DRD

-   no DRD.

The overall breakdown for discharge date for people living with frailty across all places across the full period is shown below.

```{r, echo = FALSE}

sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  summarise(Discharged_on_admission_date = sum(Discharged_on_admission_date), 
            No_DRD = sum(No_DRD), 
            Discharged_after_DRD = sum(Discharged_after_DRD), 
            Discharged_on_DRD = sum(Discharged_on_DRD), 
            Discharged_before_DRD = sum(Discharged_before_DRD)) |>
  mutate(total_discharges = Discharged_on_admission_date + 
                            No_DRD + 
                            Discharged_after_DRD  + 
                            Discharged_on_DRD + 
                            Discharged_before_DRD) |> 
  mutate(`1 Discharged on admission date` = (Discharged_on_admission_date)/total_discharges, 
         `2 No DRD` = (No_DRD)/total_discharges, 
         `5 Discharged after DRD` = (Discharged_after_DRD)/total_discharges, 
         `4 Discharged on DRD`  = (Discharged_on_DRD)/total_discharges, 
         `3 Discharged before DRD` = (Discharged_before_DRD)/total_discharges) |> 
  select(c(`1 Discharged on admission date`, 
                        `2 No DRD`, 
                        `5 Discharged after DRD`, 
                        `4 Discharged on DRD`, 
                        `3 Discharged before DRD`)) |> 
  pivot_longer(cols = c(`1 Discharged on admission date`, 
                        `2 No DRD`, 
                        `5 Discharged after DRD`, 
                        `4 Discharged on DRD`, 
                        `3 Discharged before DRD`), 
               values_to = "Percent", 
               names_to = "Discharge_Date") |> 
  transmute(`Discharge Date` = Discharge_Date, 
            Percentage = round(100* Percent, 1)) |> 
  gt()

```

The breakdown changes substantially over time, with the biggest change being a decrease in the number of discharges without a DRD (although nationally the number is still relatively high) and increases in the number of discharges on or after the DRD.

```{r, echo=FALSE}

sicb_yearmon_outcomes |> 
  subset(YearMon <= max_date) |> 
  group_by(YearMon) |> 
  summarise(Discharged_on_admission_date = sum(Discharged_on_admission_date), 
            No_DRD = sum(No_DRD), 
            Discharged_after_DRD = sum(Discharged_after_DRD), 
            Discharged_on_DRD = sum(Discharged_on_DRD), 
            Discharged_before_DRD = sum(Discharged_before_DRD)) |> 
  ungroup() |> 
  mutate(total_discharges = Discharged_on_admission_date + 
                            No_DRD + 
                            Discharged_after_DRD  + 
                            Discharged_on_DRD + 
                            Discharged_before_DRD) |> 
  mutate(`1 Discharged on admission date` = (Discharged_on_admission_date)/total_discharges, 
         `5 No DRD` = (No_DRD)/total_discharges, 
         `4 Discharged after DRD` = (Discharged_after_DRD)/total_discharges, 
         `3 Discharged on DRD`  = (Discharged_on_DRD)/total_discharges, 
         `2 Discharged before DRD` = (Discharged_before_DRD)/total_discharges) |> 
  pivot_longer(cols = c(`1 Discharged on admission date`, 
                        `5 No DRD`, 
                        `4 Discharged after DRD`, 
                        `3 Discharged on DRD`, 
                        `2 Discharged before DRD`), 
               values_to = "Percent", 
               names_to = "Discharge Date") |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = 100*Percent, 
                       fill = `Discharge Date`)) +
  geom_area(position = "stack") + 
  scale_y_continuous(name = "Frail discharges by DRD (%)") + 
  theme_minimal() + 
  theme(legend.position = "bottom")

```

Although the proportion of discharges without a DRD has come down substantially, the fact that it is still so high and changes so much during the period the programme was implemented means that it is unlikely the proportion of discharges after the DRD would be a reliable measure if used as an outcome in any impact evaluation, as it is difficult to distinguish between changes in measurement (e.g. more patients having DRDs recorded) or a change in underlying true discharge outcomes (e.g. patients being discharged more promptly after being ready to be discharged), especially as different places show different trends to each other in recording a DRD.

### Discharges to usual place of residence for people living with frailty

Discharges to usual place of residence was defined as the number of patients with a HFRS of five or higher on admission who were discharged to their usual place of residence, as a proportion of all discharged patients who had a HSF of five or higher on admission. Note that a number of patients were excluded from this analysis entirely due to having discharge location which was primarily caused by external circumstances - for example if their discharge location was a forced repatriation or was unknown.

```{r, include = FALSE}

disch <- sicb_yearmon_outcomes |> 
  transmute(sicb_code, place_name, 
            YearMon, 
            Usual_disch = admissions_equals_discharge + care_home_to_care_home + 
                          usual_residence + hospice_to_hospice, 
            Not_usual_disch = not_usual_residence + to_care_home + foster_care) |> 
  mutate(total_relevant_disch = Usual_disch + Not_usual_disch)

disch_tbl <- disch |> 
  summarise(Usual_disch = sum(Usual_disch), 
            Not_usual_disch = sum(Not_usual_disch), 
            total_relevant_disch = sum(total_relevant_disch)) |> 
  mutate(U = 100* (Usual_disch/total_relevant_disch), 
         NU = 100* Not_usual_disch/total_relevant_disch)

disch_plot <- disch |> 
  mutate(place_name = ifelse(test = is.na(place_name), 
                             yes = "Average of control places", 
                             no = place_name)) |> 
  group_by(place_name, YearMon) |> 
  summarise(Usual_disch = sum(Usual_disch), 
            Not_usual_disch = sum(Not_usual_disch)) |> 
  ungroup() |>   
  mutate(Usual.pc = Usual_disch/(Usual_disch + Not_usual_disch)) |> 
  mutate(group = ifelse(place_name == "Average of control places", 
                        yes = "Control", 
                        no = "Intervention")) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = Usual.pc, 
                       col = place_name)) + 
  geom_line() + 
  scale_y_continuous(name = "Frail discharges to usual place of residence (%)", 
                     labels = scales::label_comma()) + 
  scale_x_yearmon(name = "Date") + 
  scale_colour_discrete(name = "Place") + 
    facet_wrap(group~.,               
               nrow = 2,               
               ncol = 1) +
  theme_minimal() + 
  theme(legend.position = "bottom")
  

```

Across all places, across the whole period, `r format(round(disch_tbl$U, 1), scientific = FALSE, big.mark = ",")` % of people living with frailty were discharged to their usual place of residence.

The average percentage of people living with frailty discharged to their usual place of residence, grouped by intervention place, is shown below.

```{r, echo=FALSE}

disch |>
  subset(YearMon <= max_date) |>    
  mutate(place_name = ifelse(test = is.na(place_name),                              
                             yes = "Average of control places",             
                             no = place_name)) |>    
  group_by(place_name) |>    
  summarise(Usual_disch = sum(Usual_disch), 
            Not_usual_disch = sum(Not_usual_disch), 
            total_relevant_disch = sum(total_relevant_disch)) |> 
  ungroup() |>    
  transmute(Place = place_name,              
            'Discharged to usual place of residence (%)'= 100* round(
              Usual_disch/total_relevant_disch, 3))  |> 
    gt()
```

The national trend for the percentage of people living with frailty discharged to their usual place of residence can be seen on the chart below.

```{r, echo=FALSE}

disch |> 
  group_by(YearMon) |> 
  summarise(Usual_disch = sum(Usual_disch), 
            Not_usual_disch = sum(Not_usual_disch)) |> 
  ungroup() |> 
  mutate(Relevant_disch = Usual_disch + Not_usual_disch) |> 
  transmute(YearMon, 
         'Discharged to usual place of residence (%)' = round(
           100 * (Usual_disch/Relevant_disch), 1), 
         'Discharged to any other place' =  round(
           100* (Not_usual_disch/Relevant_disch), 1)) |> 
  ggplot(mapping = aes(x = YearMon, 
                       y = `Discharged to usual place of residence (%)`)) + 
  geom_line() + 
  scale_x_continuous("Date") + 
  theme_minimal()
```

The place level trend for the percentage of people living with frailty discharged to the usual place of residence per month is shown on the chart below (relative to the average for all the control places).

```{r, echo=FALSE}


plotly::ggplotly(disch_plot)
```
